Performance comparison of Dask and Apache Spark on HPC systems

### 2022-02-23 latex/paper.tex
[UPDATE] Modify resutls and discussion about Kmeans + minor fixes.

Figure \ref{fig:histogram_block} shows the execution time for the \textit{Histogram} when varying the number of files.
The results are similar to varying the number of nodes with Spark having faster compute time for \textit{Calculate\_histogram}.
This results in a \HL{significant} makespan difference of up to 3\%.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_histogram_block.jpg}
	\caption{Histogram: varying \# of files -- 4 nodes}
	\label{fig:histogram_block}
\end{figure}

@@

@@

\subsection{Memory management}
Our Kmeans experiment shows a significant difference between Dask and Spark.
The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
This is unexpected, given that Dask and Spark are configured similarly and \HL{are allocated} the same amount of resources.
We think those issues arise from a combination of
(1) the reserved memory for the JVM heap space,
(2) the serialization of Python to Java kept in memory,
(3) Spark only using procesess compared to Dask using threads, thus limiting shared memory consumption,
and (4) different policies to persist data to disk when running out of memory.

@@

@@

The experiments from~\cite{Mehta:17} found that Dask has a larger startup overhead than Spark, while another study~\cite{8588652} found the opposite.
In this study, we found that Dask has a lower startup overhead than Spark; however, we did not in our previous study~\cite{8943502}.
Those apparent contractions most likely come from differences in infrastructure and configuration of the engines.

@@

\section*{Acknowledgment}
Mathieu Dugr\'e was funded by an Alexander Graham Bell scholarship from
the Natural Sciences and Engineering Research Council of Canada. This work
was partially funded by the Canada Research Chairs program. The HPC cluster
was funded by the Canada Foundation for Innovation.

>>>>

Figure \ref{fig:histogram_block} shows the execution time for the \textit{Histogram} when varying the number of files.
The results are similar to varying the number of nodes with Spark having faster compute time for \textit{Calculate\_histogram}.
This results in a \HL{significant} makespan difference \HL{with Spark being up to 3\% faster}.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_histogram_block.jpg}
	\caption{Histogram: varying \# of files -- 4 nodes}
	\label{fig:histogram_block}
\end{figure}

@@

\HL{
	In a follow up experiement, we further optimized the Kmeans algorithm to minimize the size of data kept in memory.
	Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} show the gantt chart for the execution
	of this new experiments for Dask and Spark.
	The x-axis shows the time in second while the y-axis depicts the workers, with 
	one worker thread ID (Dask) and process ID (Spark) per line.
	Each task is represented on the gantt chart with the color of it's associated function.
	We observe that Spark report 3 times more process than Dask threads.
	However, we found out that both Dask and Spark use the same amount of maximum concurrent processes/threads.
	This is most likely due to the difficulty to track process id within Spark.
}

\HL{
	Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}}
	show that Dask has a significantly shorther makespan than Spark, {\SI{338}{\second}}
	and {\SI{409}{\second}} respectively.
	The main differences in makespan for Spark appear from a longer initial load of the data
	and a longer scheduling delay between the data load and the initial computation.
}

@@

\subsection{Memory management}
Our Kmeans \HL{resutls from Figure~{\ref{fig:kmeans_worker}} show} a significant difference between Dask and Spark.
The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
This is unexpected, given that Dask and Spark are configured similarly and \HL{are allocated} the same amount of resources.
We think those issues arise from a combination of
(1) the reserved memory for the JVM heap space,
(2) the serialization of Python to Java kept in memory,
(3) Spark only using procesess compared to Dask using threads, thus limiting shared memory consumption,
and (4) different policies to persist data to disk when running out of memory.

@@

\subsection{Scheduling overhead}
\HL{
	Overall, our results show that Dask has a slightly longer startup overhead compared to Spark.
	However, we found that this did not lead directly to makespan difference.
	While both engines showed similar overhead for applications with low inter-worker communication,
	Spark had significantly more scheduling overhead than Dask for the Kmeans applications.
	Spark was considerably slower to schedule the first set of computation after reading the data on the workers.
	Moreover, Spark read all the data before starting computation while was still reading data along side doing computation.
	This difference in scheduling is beneficial for data-intensive applications as it reduce the risk
	of I/O bottleneck, as explained in Section~{\ref{subsec:overhead-tradeoff}}.
}

@@

The experiments from~\cite{Mehta:17} found that Dask has a larger startup overhead than Spark, while another study~\cite{8588652} found the opposite.
In this study, we found that Dask has a \HL{larger} startup overhead than Spark; however, we did not in our previous study~\cite{8943502}.
Those apparent contractions most likely come from differences in infrastructure and configuration of the engines.

@@

\section*{Acknowledgment}
Mathieu Dugr\'e was funded by \HL{the} Alexander Graham Bell scholarship from
the Natural Sciences and Engineering Research Council of Canada. This work
was partially funded by the Canada Research Chairs program. The HPC cluster
was funded by the Canada Foundation for Innovation.
